Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario

### 2017-01-17 documents/TriangleScenario/ts.tex
finished 437 report

% === All of the Packages I use frequently ===
\usepackage{../../packages/document_config}
\usepackage{../../packages/shared}
\usepackage{../../packages/misc_commands}

@@

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        Quantum correlations are often incompatible with a classical assumption of causal structure. This nonclassicality is often known as quantum nonlocality, and it is witnessed through the violation of causal compatibility inequalities, such as Bell inequalities. Such inequalities were recently derived for the Triangle scenario [arXiv:1609.00672], begging the question: can these inequalities be violated by quantum correlations? Here we answer this affirmatively, and discuss specific Triangle scenario inequalities and quantum configurations which manifest nonclassical correlations. Numerical optimzations reveal quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents

    \section{Introduction}
    \begin{itemize}
        \item \todo[TC]{Overview of importance of inequalities}
        \item \todo[TC]{Triangle Scenario and existing work}
        \item \todo[TC]{Objective of research project}
        \item \todo[TC]{Structure of this paper}
    \end{itemize}

    \section{Causal Compatibility}
    \begin{itemize}
        \item \todo[TC]{Define marginal scenario}
        \item \todo[TC]{Define marginal model}
        \item \todo[TC]{Define causal structure}
        \item \todo[TC]{Define compatibility}
    \end{itemize}

    \section{Triangle Scenario}
    \begin{itemize}
        \item \todo[TC]{Discuss some of its appearances in other work}
        \item \todo[TC]{Figure}
    \end{itemize}
    \subsection{Fritz Distribution}
    \label{sec:fritz_distribution}
    \begin{figure}
    \begin{center}
        % \fbox{%
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/triangle_scenario_operational_fritz_bell}}
            \caption{The triangle scenario re-imagined to mimic the Bell scenario. The measurement settings $S_{\p{A}},S_{\p{B}}$ are latent nodes unlike the Bell scenario (\cref{fig:bell_scenario}).}
            \label{fig:triangle_scenario_with_fritz_bell_embedded}
        \end{minipage}%
        % }%
        \hspace{0.04\textwidth}%
        % \fbox{%
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/bell_scenario}}
            \caption{The Bell scenario consisting of two observers $\p{A}, \p{B}$ together with measurement settings $S_{\p{A}}$ and $S_{\p{B}}$ respectively.}
            \label{fig:bell_scenario}
        \end{minipage}%
        % }%
    \end{center}
    \end{figure}
    As was originally noticed by Fritz~\cite{Fritz_2012}, it is possible to construct quantum distributions incompatible with the triangle scenario by utilizing quantum distributions incompatible with the familiar Bell scenario. To explain, imagine rearranging the triangle into the configuration depicted in \cref{fig:triangle_scenario_with_fritz_bell_embedded} and contrast it with the Bell scenario of \cref{fig:bell_scenario}. The crucial distinction to be made is that the respective measurement settings $S_{\p{A}}, S_{\p{B}}$ in the Bell scenario become latent nodes in the triangle scenario. In order to embed Bell scenario incompatibility into the triangle scenario, the latent nodes $S_{\p{A}}, S_{\p{B}}$ need to be observable and independent of the shared latent node $\la$. This can be accomplished by having $C$ measure the measurement settings $S_{\p{A}}, S_{\p{B}}$ and announce them as an outcome. Consequently, any distribution over $\p{A}, \p{B}, S_{\p{A}}, S_{\p{B}}$ that is incompatible with Bell scenario is also incompatible with the triangle scenario provided that $C$ is perfectly correlated with $S_{\p{A}}, S_{\p{B}}$. Understandably there are numerous ways that this can be accomplished, albeit we will focus on elucidating an exemplary case.

    The \term{Fritz distribution}, denoted $\prob[\fritz]$, is a quantum-accessible distribution known to be incompatible with the triangle scenario~\cite{Fritz_2012}. Explicitly, the Fritz distribution can be written as follows:

@@

            \includegraphics[scale=0.6]{../../figures/distributions/fritz_dist_plot_bits_no_title.pdf}

@@

    Consequently, it is possible to define the correlation between right-bits of $A$ and $B$,

@@

    And also define a CHSH inequality~\cite{CHSH_Original} for the right-bits of $A$ and $B$,

@@

    Substitution of \cref{eq:fritz_dist} into \cref{eq:CHSH} yields maximal violation~\cite{Cirelson_1980},

@@

    The Fritz distribution $\prob[\fritz]$ can be realized as a set of quantum states and measurements. The shared resource between $A$ and $B$ (graphically denoted $\la$) is the maximally entangled Bell state:
    \[ \ket{\Phi^+} = \f{1}{\sqrt{2}}\br{\ket{00} + \ket{11}} \qquad \rho_{AB} = \ket{\Phi^+}\bra{\Phi^+} \]
    While the states shared with $C$ are the following mixed states:
    \[ \rho_{BC} = \rho_{CA} = \f{\ket{00}\bra{00} + \ket{11}\bra{11}}{2} \]
    The measurements employed by $A,B$ and $C$ are separable, projective measurements,
        M_{A} &= \bc{\ket{0\psi_{1}}\bra{0\psi_{1}}, \ket{0\psi_{5}}\bra{0\psi_{5}}, \ket{1\psi_{3}}\bra{1\psi_{3}}, \ket{1\psi_{7}}\bra{1\psi_{7}}} \\
        M_{B} &= \bc{\ket{\psi_{6}0}\bra{\psi_{6}0}, \ket{\psi_{2}0}\bra{\psi_{2}0}, \ket{\psi_{0}1}\bra{\psi_{0}1}, \ket{\psi_{4}1}\bra{\psi_{4}1}} \\
        M_{C} &= \bc{\ket{00}\bra{00}, \ket{10}\bra{10}, \ket{01}\bra{01}, \ket{11}\bra{11}}
    Where $\ket{\psi_n}$ is shorthand for,
    \[ \ket{\psi_n} = \f{1}{\sqrt{2}}\br{\ket{0} + e^{in/4}\ket{1}} \]

    Before continuing it is worth noting that \cref{eq:fritz_dist} is non-unique. Any distribution that is equal to \cref{eq:fritz_dist} via a permutation of outcomes or exchange of parties is should also be referred to as a Fritz distribution. Moreover, the quantum realization presented here is not unique either. Nonetheless for concreteness, \cref{eq:fritz_dist} is taken as \textit{the} Fritz distribution throughout this paper.

    Additionally, it is important to understand the domain in which Fritz's proof of incompatibility is valid. Specifically, \cref{eq:CHSH} can only be applied to the triangle scenario where there exists perfect correlation between $\p C$'s outcomes and the measurement pseudo-settings (left-bits) of $\p A$ and $\p B$. The details of this restriction are discussed and proven in Fritz's original work~\cite{Fritz_2012}. As such, the incompatibility proof employed by Fritz is not applicable to any distribution that significantly deviates from the Fritz distribution presented above. For example, if one affinely combines \cref{eq:fritz_dist} with uniform noise, at what point does the resultant distribution become compatibility? This problem is discussed more in \cref{sec:noise}.

>>>>

% === All of the Packages I use frequently ===
\usepackage{../../packages/document_config}
\usepackage{../../packages/shared}
\usepackage{../../packages/misc_commands}
\usepackage{subcaption}
\usepackage{enumitem}

\renewcommand{\Events}[1]{\mathcal{S}\br{#1}}
\renewcommand{\tcdot}{\cdot}

@@

\begin{document}
    \title{Causal Compatibility Inequalities Admitting of Quantum Violations in the Triangle Scenario}
    \author{Thomas C. Fraser}
    \email{tcfraser@tcfraser.com}
    \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada \\ University of Waterloo, Waterloo, Ontario, Canada}
    % \author{Elie Wolfe}
    % \email{ewolfe@perimeter@institute.ca}
    % \affiliation{Perimeter Institute for Theoretical Physics, Waterloo, Ontario, Canada}
    \date{\today}
    \begin{abstract}
        Quantum correlations are often incompatible with a classical assumption of causal structure. This non-classicality is often known as quantum non-locality, and it is witnessed through the violation of causal compatibility inequalities, such as Bell inequalities. Such inequalities were recently derived for the triangle scenario [arXiv:1609.00672], begging the question: can these inequalities be violated by quantum correlations? Here we answer this affirmatively, and discuss specific triangle scenario inequalities and quantum configurations which manifest non-classical correlations. Numerical optimizations reveal quantum resources potentially qualitatively different from those known previously.
    \end{abstract}
    \maketitle
    \tableofcontents
    \clearpage

    \section{Introduction}
    \label{sec:introduction}
    In recent decades, the technical utility of quantum mechanics has become abundantly clear; quantum computational algorithms such as Shor's algorithm~\cite{Shor_1997} and numerous others~\cite{Jordan_2016} maintain advantage over their classical counterparts. Additionally, classical communications technologies stand to benefit from the nature of quantum mechanics --- a popular example being Quantum Key Distribution~\cite{Bennett_2014}. Any time classical resources are unable to emulate quantum mechanics, there represents an opportunity to exploit the non-classical features of quantum mechanics in order to solve a variety of computational and communicational problems~\cite{Neilsen_Chaung_2011}. Albeit, not all tasks stand to benefit from quantum resources~\cite{Almeida_2010}. Presently, discovering situations or tasks where quantum mechanics offers an advantage, and proving that advantage is genuine, is a primal objective of theoretical quantum information theory. \\

    From a foundational prospective, the majority of attempts to characterize precisely where quantum mechanics and classical mechanics differ have involved Bell inequalities~\cite{Bell_1964}. Originally, Bell inequalities were derived in response to the famous EPR paradox~\cite{EPR_Orig} to rule out hidden variable theories for quantum mechanics. Studying and discovering Bell-type inequalities is useful because a complete set of Bell inequalities provides a systematic way of classifying the \textit{classicality} of a given observation. These inequalities are best understood from the perspective of \term{causal inference} --- a problem which aims to determine which observations can or can not be explained by a hypothesized causal structure. The abstract nature of causal inference is responsible for its presence in numerous scientific fields including machine learning and biology~\cite{Pearl_2009,Pearl_2009_tr}. Inequalities such as Bell inequalities constrain the space of observations that are compatible with a hypothesized causal structure. In the case of the original Bell inequalities, the classical causal model in question is known as the Bell scenario~\cite{Wood_2012} and is depicted in \cref{fig:bell_scenario}. The Bell scenario involves two parties ($A, B$) who each make measurements decided by random settings ($S_A, S_B$) on a shared hidden resource $\la$. Quantum non-classicality in the Bell scenario and related structures have been studied thoroughly since Bell's original work~\cite{Brunner_2013}. However, novel scenarios such as the correlation scenarios proposed by Fritz~\cite{Fritz_2012,Fritz_2014} are arguably less understood compared to the Bell scenario. Here we investigate one particular causal scenario: the triangle scenario (\cref{fig:triangle_scenario}). \\

    The triangle scenario (\cref{fig:triangle_scenario}) is a causal structure composed of $3$ parties labeled $A, B, C$ arranged in a triangular configuration while pair-wise sharing hidden/latent variables $X, Y, Z$. It has been studied extensively in literature before (see~\cite[Fig. 1]{Steudel_2010},~\cite[Fig. 6]{Chaves_2014},~\cite[Fig. 8]{Branciard_2012},~\cite[Fig. 8, Appendix E]{Henson_2014},~\cite[Fig. 3]{Fritz_2012},~\cite[Fig. 1]{Inflation}, $\ldots$). A healthy portion of which the existing literature will be discussed in \cref{sec:triangle_scenario}. For instance \citet{Branciard_2012} noted that characterizing locality in the triangle scenario remained an \textit{open problem} and that identifying causal compatibility inequalities for this configuration \textit{seemed challenging}. The subject of this document is to expound on the developments made towards understanding quantum non-classicality in the triangle scenario. \\

    As a preliminary summary, causal compatibility inequalities were derived for the triangle scenario that are able to be violated by specific quantum-accessible distributions. This was accomplished by combining two previous works. The first, an insight made by \citet{Fritz_2012} regarding the ability to re-interpret the Bell scenario as a portion of the triangle scenario. The second, a new framework for solving causal inference problems developed by Wolfe \emph{et al.}~\cite{Inflation} called \term{The Inflation Technique}. Ultimately, this work serves as a strong indication that the Inflation Technique offers new insights into specifically quantum non-classicality. Moreover, these new inequalities offer an avenue for finding and identifying new forms of non-classicality unknown previously. This report discusses attempts to find such resources with partial success and outlines refinements that can be made for future exploration.\\

    The following paper is organized into the following sections: \Cref{sec:causal_compatibility} recalls important notions from causal inference theory and sets up the notation to be used. \Cref{sec:triangle_scenario} discusses the triangle scenario and its existing research in detail; identifying its stark differences from the Bell scenario and motivates why the triangle scenario was chosen as a case study. \Cref{sec:fritz_distribution} defines and discusses a singular quantum correlation first discovered by \citet{Fritz_2012} which we term the \term{Fritz distribution}. The Fritz distribution was proven to be non-classical without the use of inequalities --- this report derives inequalities that do. \Cref{sec:inflation_technique} briefly summaries the Inflation Technique in the context of this work. Although the summary presented here is sufficient, the original work~\cite{Inflation} offers a more concrete and pedagogical introduction. \Cref{sec:deriving_inequalities} demonstrates how the inflation technique can be used to derive causal compatibility inequalities for the triangle scenario that are violated by quantum-accessible distributions. A sample of these inequalities are presented at the end of \cref{sec:deriving_inequalities} as well as \cref{sec:web_inequality,sec:symmetic_web_inequality}. Finally, \cref{sec:violations_noise} mentions a method for parameterizing the space of quantum-accessible distributions which is used to numerically optimize against our derived inequalities in search of stronger types of non-classicality. Additionally, the Fritz distribution is subjected to noise in order to measure the robustness of the derived inequalities.

    \section{Causal Compatibility}
    \label{sec:causal_compatibility}
    The task of causal inference aims to determine whether or not a given family of observed probability distributions can be explained by a proposed causal mechanism~\cite{Pearl_2009}. If a set of observed distributions can be explained by the hypothesized causal mechanism, then the observations are said to be \textit{compatible} with the causal mechanism. In order to define compatibility rigorously, we first need to formally define what is meant by ``set of observations''.

    \subsection{Marginal Models}
    Given a set of observable variables $\jointvar = \bc{v_1, \ldots, v_k}$ where each $v \in \jointvar$ is a statistical random variable, one can conceivably make a measurement over of these variables and obtain a joint probability distribution $\prob[\jointvar] = \prob[v_1,\ldots,v_n]$. More generally however, and depending on the context, it is possible to make a measurement on a portion of the system $V \subset \jointvar$ in order to obtain a marginal distribution $\prob[V]$. With complete generality however, one could in principle make a measurement on a collection of portions of the system $\mscenario = \bc{V_1, \ldots, V_m \mid \forall i : V_i \subseteq \jointvar}$ obtaining a collection of marginal probability distributions $\bc{\prob[V_1], \ldots, \prob[V_m] \mid \forall i: V_i \in \mscenario}$. \\

    The set $\mscenario = \bc{V \mid V \subseteq \jointvar}$ of subsets of $\jointvar$ is referred to as the \term{marginal scenario} and each element $V \in \mscenario$ is termed a \term{(marginal) context} of $\mscenario$. The complete set of marginal distributions is referred to as the \term{marginal model} and is denoted with an superscript $\prob^{\mscenario}$:
    \[ \prob^{\mscenario} = \bc{\prob[V] \mid V \in \mscenario} \]
    A marginal model acts as the most general description of a family of observations that can be made over $\jointvar$. Strictly speaking, as defined by~\cite{Fritz_2011}, a marginal scenario forms an \textit{abstract simplicial complex} where it is required that all subsets of contexts are also contexts.
    \[ \forall V \in \mscenario, V' \subset V : V' \in \mscenario \]
    Throughout this document, we only consider (without loss of generality) the \textit{maximal} marginal scenario; restricting our focus to the largest contexts in the marginal scenario. Additionally, all marginal scenarios are taken to be \text{complete} in the sense that the marginal scenario covers the complete set of observable variables $\jointvar$.
    \[ \jointvar = \bigcup_{V \in \mscenario} V \]

    \subsection{Causal Structure}
    A \term{causal structure} abstractly represents the hypothesized causal mechanism. Mathematically, a causal structure is a directed acyclic graph $\graph = \br{\nodes, \edges}$ of nodes $\nodes$ and edges $\edges$ as defined in \cref{sec:graph_theory}. The nodes of a causal structure are random variables and are classified into two disjoint classes: the \term{latent nodes} $\nodes_L$ representing the set of nodes that are unobservable and the \term{observed nodes} $\nodes_O$. The latent nodes of a causal structure represent completely hidden variables that are hidden by some fundamental process or cannot be measured due to other limitations. The observed nodes represent random variables that are measurable and in practice will be observed by the marginal model $\prob^\mscenario$. The edges $\edges$ of a causal structure represent casual influence between variables within the causal structure. \\

    A number of causal structures are depicted throughout this document (see \cref{fig:bell_scenario,fig:triangle_scenario} as well as \cref{fig:inflations}). As an example, the triangle scenario has $3$ observable nodes and $3$ latent nodes,
    \[ \nodes_{O} = \bc{A, B, C} \qquad \nodes_{L} = \bc{X, Y, Z} \]
    along with $6$ edges indicating that the latent nodes $\bc{X, Y, Z}$ cause and influence the observed nodes $\bc{A, B, C}$.
    \[ \edges = \bc{\bc{X \to A}, \bc{Y \to A}, \bc{Y \to B}, \bc{Z \to B}, \bc{Z \to C}, \bc{X \to C}} \]

    \subsection{Compatibility}
    \label{sec:compatibility}

    Now that the notions of a causal structure $\graph$ and a marginal model $\prob^{\mscenario}$ are defined, it is now possible to precisely define the notion of compatibility between the two. Intuitively, the causal connections within $\graph$ should restrict and limit the set of possible observations that can be made on $\graph$. Namely, if each variable $n \in \nodes$ is caused by its parents $\Pa[\graph]{n}$, then $n$ should be independent of all other variables that are not its descendants. This is known as the \term{local Markov property}:
    \[ n \indep \nodes \setminus \De[\graph]{n} \mid \Pa[\graph]{n} \quad \forall n \in \nodes \eq \label{eq:local_markov_property} \]
    Where $\indep$ denotes statistical independence. With this as motivation, we define a set of \term{causal parameters} for a causal structure $\graph$ to be a specification of a conditional distribution for every variable $n \in \nodes$ given its parents:
    \[ \bc{\prob[n \mid \Pa[\graph]{n}] \mid n \in \nodes} \eq \label{eq:causal_parameters} \]
    It is important to note that a set of causal parameters is not unique; the numerical nature of set of causal parameters is a \text{choice}. Altogether, a marginal model $\prob^{\mscenario}$ is \term{compatible} with a causal structure $\graph$ if there exists a choice of causal parameters such that each context $\prob[V] \in \prob^{\mscenario}$ can be \textit{recovered} from the following series of operations:
    \begin{enumerate}
        \item Obtain a joint probability distribution over \textit{all} variables of the causal structure using the chosen casual parameters:\footnote{Here $\prod$ denotes the product distribution $\prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] = \prob[n_1 \mid \Pa[\graph]{n_1}] \times \cdots \times \prob[n_k \mid \Pa[\graph]{n_k}]$.}
        \[ \prob[\nodes] = \prod_{n \in \nodes} \prob[n \mid \Pa[\graph]{n}] \eq \label{eq:local_markov_property_operation}\]
        This step ensures that the local Markov property (\cref{eq:causal_parameters}) holds for all variables $n \in \nodes$.
        \item Marginalize over the latent variables of $\graph$ to obtain a joint distribution over the observed variables $\nodes_O$:
        \[ \prob[\nodes_O] = \sum_{\nodes_L} \prob[\nodes] \eq \label{eq:observable_marginalization} \]
        \item Finally marginalize over the observed nodes not in $V$ to obtain $\prob[V]$:
        \[ \prob[V] = \sum_{\nodes_O \setminus V} \prob[\nodes_O] \eq \label{eq:context_marginalization} \]
    \end{enumerate}
    It is important to note that compatibility between $\prob^{\mscenario}$ and $\graph$ requires that it is possible to recover \textit{each} context $\prob[V] \in \prob^{\mscenario}$ from the \textit{same} set of causal parameters. If it is not possible for $\prob^{\mscenario}$ to be recovered for \textit{any} choice of causal parameters, then $\prob^{\mscenario}$ is \term{incompatible}.

    \subsection{Inequalities}
    As defined in \cref{sec:compatibility}, determining compatibility is equivalent to determining if a set of causal parameters exists with the above-mentioned properties. In practice however, checking causal parameters systemically is intractable, especially when the cardinality (number of outcomes) of the latent variables is unspecified. Instead, \term{causal compatibility inequalities}\footnote{We refer to these inequalities as ``causal compatibility inequalities'' instead of simply Bell inequalities for two reasons. First, ``Bell inequalities'' usually are associated specifically with the Bell scenario/Bell causal structure. Second, the inequalities derived in this work are fundamentally distinct from a typical Bell inequality in that these inequalities are \textit{polynomial} over $\prob^{\mscenario}$ instead of \textit{linear}.} are probabilistic inequalities defined over the marginal contexts $\prob[V] \in \prob^{\mscenario}$ within the marginal model. If a marginal model $\prob^{\mscenario}$ happens to \textit{violate} any causal compatibility inequality, then that marginal model is definitely \textit{incompatible}. In \cref{sec:deriving_inequalities}, a brief discussion of existing techniques for deriving these inequalities will take place.\\

    Unfortunately, a singular inequality can not prove that a given marginal model is \textit{compatible}, only \textit{incompatible}. Instead, a \textit{complete characterization} of compatibility would consist of a complete set of all valid causal compatibility inequalities; satisfaction of a complete set of inequalities does prove compatibility. Currently however, it is unknown how to obtain a complete characterization for any given causal structure, including the triangle scenario.  \\

    From the perspective of identifying quantum non-classicality, the causal structure $\graph$ acts as the classical hypothesis for a particular marginal model $\prob^{\mscenario}$. Therefore, non-classicality becomes synonymous with incompatibility: if a marginal model $\prob^{\mscenario}$ is incompatible, then it is non-classical. Henceforth, we will use these two terms interchangeably. From a resource standpoint, if the non-classicality of $\prob^{\mscenario}$ can be witnessed by an inequality $I$, but $\prob^{\mscenario}$ can be implemented using quantum states and measurements, then the causal compatibility inequality $I$ represents a concrete task where quantum resources out-perform classical resources.

    \section{Triangle Scenario}
    \label{sec:triangle_scenario}

    \begin{figure}
    \begin{center}
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/bell_scenario}}
            \caption{The Bell scenario consisting of two observers $\p{A}, \p{B}$ together with measurement settings $S_{\p{A}}$ and $S_{\p{B}}$ respectively. The shared hidden variable is labeled $\la$.}
            \label{fig:bell_scenario}
        \end{minipage}\hspace{0.04\textwidth}%
        \begin{minipage}[t]{.48\textwidth}
            \centering
            \scalebox{1.0}{\input{../../figures/causal_structures/triangle_scenario}}
            \caption{The casual structure of the triangle scenario. Three variables $A,B,C$ are observable, while $X, Y, Z$ are latent variables.}
            \label{fig:triangle_scenario}
        \end{minipage}
    \end{center}
    \end{figure}
    Discovering quantum non-classicality in any causal structure is desirable, so why, in particular, is it important to do so in the triangle scenario? As was mentioned in \cref{sec:introduction,sec:causal_compatibility}, the \term{triangle scenario} (\cref{fig:triangle_scenario}) is a causal structure $\graph$ consisting of $3$ observable variables $A, B, C$ arranged in a triangular configuration while pair-wise sharing latent variables $X, Y, Z$. \\

    Following the definition of causal compatibility from \cref{sec:causal_compatibility}, a distribution $\prob[\nodes_{O}] = \prob[ABC]$ is compatible with the triangle scenario if and only if there exists a choice of causal parameters $\bc{\prob[A|X,Y],\prob[B|Y,Z],\prob[C|Z,X],\prob[X],\prob[Y],\prob[Z]}$ such that the joint distribution $\prob[\nodes] = \prob[ABCXYZ]$ is a product over the causal parameters,
    \[ \prob[ABCXYZ] = \prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \]
    And also, the given distribution $\prob[ABC]$ is a marginalization of $\prob[ABCXYZ]$ over $X, Y, Z$:
    \[ \prob[ABC] = \sum_{X,Y,Z}\prob[A|X,Y]\prob[B|Y,Z]\prob[C|Z,X]\prob[X]\prob[Y]\prob[Z] \eq \label{eq:triangle_compatibility} \]

    For certain causal structures, the conditional independence relations provided by $d$-separation relations are a sufficient characterization of compatibility. Recently, \citet{Henson_2014}, classified the triangle scenario as an \textit{interesting} causal structure in the sense that conditional independence relations \textit{do not} provide a sufficient characterization of non-classicality.\footnote{Note that the triangle scenario possesses \textit{zero} conditional independence relations over the observable nodes $A, B, C$.} Moreover, \citet{Inflation} concretely demonstrate that the space classical distributions on the triangle scenario is non-convex, unlike the Bell scenario. It can be argued that convexity of the Bell scenario distributions is partially responsible for the wealth of knowledge that is known about the Bell scenario. It is also reasonable to speculate that quantum non-classicality in the triangle can be translated directly into a computational advantage for certain computational circuits~\cite{Terhal_2002}. Characterizing quantum non-classicality for interesting causal structures remains an unsolved problem, and the triangle scenario is a small, non-trivial case study. \\

    Additionally, \citet{Fritz_2012} demonstrated that the triangle scenario is the \textit{smallest} correlation scenario\footnote{Correlation scenarios differ from the Bell scenario in that measure setting variables are non-existence in correlation scenarios.} in which their exists incompatible, quantum distributions by explicitly mentioning one which we have elected to call the Fritz distribution (see \cref{sec:fritz_distribution}). Fritz's proof of non-classicality does not involve compatibility inequalities, but instead mentions that inequalities would be helpful for characterizing non-classical in the triangle scenario. \\

    Finally \citet{Inflation} proposed the Inflation Technique (summarized in \cref{sec:inflation_technique}), a novel tool for solving causal inference problems. Using the inflation technique, they were able to prove incompatibility between the triangle scenario and the $W$-type distribution; a quantum non-accessible distribution whose incompatibility is not witness-able by \textit{any} entropic inequalities or any other known constraints. Therefore as a an algorithmic framework, the Inflation Technique comes closer to characterizing non-classicality in the triangle scenario that all others. \\

    It is worth noting that a number of causal compatibility inequalities for the triangle scenario have been derived in previous works. For example, \citet{Steudel_2010} derived an inequality for a series of causal structures with $n \in \N$ variables where at most $c \in \N$ share a common ancestor. The triangle scenario is a special case with $c = 2$, $n = 3$. Moreover, \citet{Henson_2014} derive a family of entropic inequalities constraining the set of compatible correlations. Most impressively, \citet{Inflation} derive a family of polynomial, $2$-outcome, causal compatibility inequalities; $37$ non-trivial representatives are printed in~\cite{Inflation}. \\

    As a preliminary search for quantum incompatibility in the triangle scenario, we performed numerical optimizations using bipartite qubit density matrices and $2$-outcome POVM measurements against the compatibility inequalities printed in~\cite{Inflation} as well as the entropic inequalities presented by \citet{Henson_2014}. Unfortunately, none of these inequalities could be violated. These early results suggest, although do not prove, that quantum non-classical does not exist in the triangle scenario for \textit{two-outcome} measurements. We make \textit{no claim} that this is a universal truth and simply take it as motivation to explore incompatibility in triangle scenario for a degree of outcomes. \\

    In conclusion, the triangle scenario is a causal structure whose non-classicality remains an open problem. The triangle scenario is a desirable case study because there is a quantum distribution that is known to be incompatible, but no previously known inequalities were able to witness its incompatibility. This failure represents a gap in our understanding of quantum non-classicality, where recent developments (namely the Inflation Technique) have made partial progress. This document expands on our understanding of non-classicality in the triangle scenario and also our understanding on how to find inequalities using the Inflation Technique for interesting causal structures in general.

    \section{Fritz Distribution}
    \label{sec:fritz_distribution}
    \begin{figure}
    \begin{center}
        \scalebox{1.0}{\input{../../figures/causal_structures/triangle_scenario_operational_fritz_bell}}
        \caption{The triangle scenario re-imagined to mimic the Bell scenario. The measurement settings $S_{\p{A}},S_{\p{B}}$ are latent nodes unlike the Bell scenario (\cref{fig:bell_scenario}).}
        \label{fig:triangle_scenario_with_fritz_bell_embedded}
    \end{center}
    \end{figure}

    As was originally noticed by \citet{Fritz_2012}, it is possible to construct quantum distributions incompatible with the triangle scenario by utilizing quantum distributions incompatible with the familiar Bell scenario. To explain, imagine rearranging the triangle into the configuration depicted in \cref{fig:triangle_scenario_with_fritz_bell_embedded} and contrast it with the Bell scenario of \cref{fig:bell_scenario}. Evidently, under the correct relabeling, large portions of the Triangle Scenario resemble the Bell Scenario. The crucial distinction is that the measurement settings $S_{\p{A}}, S_{\p{B}}$ are \textit{observable variables} in the Bell scenario but become \textit{latent variables} in the triangle scenario. In order to embed distributions that are incompatible with the Bell scenario into the triangle scenario, the latent nodes $S_{\p{A}}, S_{\p{B}}$ need to be observable and independent of the shared latent node $\la$. This can be accomplished by having $C$ ``measure'' the measurement settings $S_{\p{A}}, S_{\p{B}}$ and announce them as an outcome. Consequently, any distribution over $\p{A}, \p{B}, S_{\p{A}}, S_{\p{B}}$ that is incompatible with Bell scenario is also incompatible with the triangle scenario provided that $C$ is \textit{perfectly correlated} with $S_{\p{A}}, S_{\p{B}}$. Understandably there are numerous ways that this can be accomplished, albeit we will focus on elucidating an exemplary method. \\

    The \term{Fritz distribution}, denoted $\prob[\fritz]$, is a quantum-accessible distribution known to be incompatible with the triangle scenario~\cite{Fritz_2012}. In the Fritz distribution, each of the variables $A,B,C$ have $4$ possible outcomes. Explicitly, $\prob[\fritz]$ can be written as:

@@

            \includegraphics[scale=0.6]{../../figures/distributions/fritz_dist_plotted_bits.pdf}
            \vspace{-0.2in}
            \[ \probplotvalue{253, 253, 84} = \f{1}{32}\br{2 - \sqrt{2}} \qquad \probplotvalue{11, 1, 0} = \f{1}{32}\br{2 + \sqrt{2}}\]

@@

    Consequently, it is possible to define the correlation between the right-hand bits of $A$ and $B$ as the probability that $A_r$ and $B_r$ are equal less the probability that they differ:

@@

    Provided that $C$ is perfectly correlated with $A_l$ and $B_l$, any Bell inequality for the Bell scenario defined over $\p{A}, \p{B}, S_{\p{A}}, S_{\p{B}}$ can be directly converted to an inequality for the triangle scenario using the following mapping:
    \[ \p{A}, \p{B}, S_{\p{A}}, S_{\p{B}} \mapsto \p{A_r}, \p{B_r}, \p A_l, \p B_l \]
    As an example, the famous CHSH inequality~\cite{CHSH_Original},
    \[ \ba{\p A\p B | S_{\p{A}} = 0, S_{\p{B}} = 0} + \ba{\p A\p B | S_{\p{A}} = 0, S_{\p{B}} = 1} + \ba{\p A\p B | S_{\p{A}} = 1, S_{\p{B}} = 0} - \ba{\p A\p B | S_{\p{A}} = 1, S_{\p{B}} = 1} \leq 2 \]
    Can be written for the right-bits of $A$ and $B$,

@@

    Therefore, every compatible distribution $\prob[ABC]$ where $C$ is perfectly correlated $A_l,B_l$ must satisfy \cref{eq:CHSH}. Substituting the Fritz distribution (\cref{eq:fritz_dist}) into \cref{eq:CHSH} yields maximal violation~\cite{Cirelson_1980},

@@

    Therefore, the Fritz distribution is incompatible with the triangle scenario as promised. \\

    The Fritz distribution $\prob[\fritz]$ is of special interest because it can be realized as a set of quantum states and measurements on the quantum causal model depicted in \cref{fig:triangle_scenario_quantum_model}. The shared resource between $A$ and $B$ (graphically denoted $\la$ in \cref{fig:triangle_scenario_with_fritz_bell_embedded}) is the maximally entangled Bell state:
    \[ \ket{\Phi^+} = \f{1}{\sqrt{2}}\br{\ket{00} + \ket{11}} \qquad \rho_{AB} = \ket{\Phi^+}\bra{\Phi^+} \]
    While the states shared with $C$ are the following mixed states:
    \[ \rho_{BC} = \rho_{CA} = \f{\ket{00}\bra{00} + \ket{11}\bra{11}}{2} \]
    The measurements employed by $A,B$ and $C$ are separable, projective measurements,
    \begin{align*}
        M_{A} &= \bc{\ket{0\psi_{1}}\bra{0\psi_{1}}, \ket{0\psi_{5}}\bra{0\psi_{5}}, \ket{1\psi_{3}}\bra{1\psi_{3}}, \ket{1\psi_{7}}\bra{1\psi_{7}}} \\
        M_{B} &= \bc{\ket{\psi_{6}0}\bra{\psi_{6}0}, \ket{\psi_{2}0}\bra{\psi_{2}0}, \ket{\psi_{0}1}\bra{\psi_{0}1}, \ket{\psi_{4}1}\bra{\psi_{4}1}} \\
        M_{C} &= \bc{\ket{00}\bra{00}, \ket{10}\bra{10}, \ket{01}\bra{01}, \ket{11}\bra{11}}
    \end{align*}
    Where $\ket{\psi_n}$ is shorthand for,
    \[ \ket{\psi_n} = \f{1}{\sqrt{2}}\br{\ket{0} + e^{in/4}\ket{1}} \]
    Using the above states and measurements, the Fritz distribution $\prob[\fritz]$ can be recovered using the following:
    \[ \prob[ABC] = \Tr\bs{\netperm^\intercal \rho_{AB}\otimes\rho_{BC}\otimes\rho_{CA} \netperm M_{A}\otimes M_{B} \otimes M_{C}} \]
    Where $\netperm$ is a $64 \times 64$ permutation matrix to be elaborated in \cref{sec:perm_matrix}. \\
